### 2020-01-17 ms/ms.tex
Incorporate Brian's comments and improve discussion on potential mechanisms of climate effects

\usepackage[round,sectionbib]{natbib}
\setcitestyle{authoryear}
\bibpunct{(}{)}{;}{a}{}{;}
\bibliographystyle{mee}

@@

\noindent\small{\textsuperscript{1} Department of Biology, Dalhousie University, Halifax, NS, B3H 4R2, Canada\\}
\small{\textsuperscript{2} Norwegian Institute for Nature Research (NINA), Trondheim, Norway\\}
\small{\textsuperscript{3} Department of Fisheries and Oceans Canada, St. John's, NL, Canada\\}
\small{\textsuperscript{4} Minist\`{e}re des For\^{e}ts, de la Faune et des Parcs, Qu\`{e}bec, QC, Canada\\}
\small{\textsuperscript{5} Department of Fisheries and Oceans Canada, Moncton, NB, Canada\\}
\small{\textsuperscript{6} Department of Fisheries and Oceans Canada, Dartmouth, NS, Canada\\}
\small{\textsuperscript{*} Corresponding author: spardo@dal.ca}

@@

The marine phase of anadromous Atlantic salmon (\emph{Salmo salar}) is the
least known yet one of the most crucial with regards to population
persistence. Recently, declines in many Atlantic salmon populations in eastern
Canada have been attributed to changes in conditions at sea, negatively
affecting their survival. However, marine survival estimates are difficult to
obtain given that many individuals spend multiple winters in the ocean before
returning to freshwater to spawn, necessitating the estimation of multiple
parameters. To do so, we fit a hierarchical Bayesian maturity schedule model
to smolt and adult abundance time series for seven populations located in
Newfoundland and Labrador (NL), Qu\'{e}bec (QC), New Brunswick (NB), and Nova
Scotia (NS). The datasets ranged between 19 and 42 years in length. We
estimated three components of marine survival: survival in the first year at
sea (\So), survival in the second year at sea (\St), and proportion returning
as one sea-winter adults (\Pg). Controlling for time frame, trends in
estimates of \So were not consistent among rivers.
In the four populations for which data extended to 1990, marine survival
during the first year at sea predominantly increased over time in Western Arm
Brook (NL) and Rivi\`{e}re Saint-Jean (QC) but largely declined in Conne River
(NL) and Rivi\`{e}re Trinit\'{e} (QC). In the three other populations, \So
exhibited an increase in Campbellton River (NL), a decline in LaHave River
(NS), and fluctuating stability in Nashwaak River (NB) since approximately the
mid-1990s. 
Estimates of \St were highly uncertain and trends could not be assessed. \Pg
was temporally stable, except for LaHave and Nashwaak Rivers. Our findings
challenge the assumption that marine survival has changed in a consistent
manner among populations across a broad geographic scale, and suggest that
trends can be river-specific. Further work exploring potential correlates of
these trends in marine survival is warranted.

@@

Reductions in fishing mortality, albeit necessary, are not always sufficient
to facilitate population recovery. Experience with numerous commercially
exploited marine fisheries since the early 1990s has shown that not all
populations respond as favourably as anticipated to considerable reductions in
exploitation \citep{Hutchings2017}. Efforts to close commercial
Atlantic salmon (\emph{Salmo salar}) fisheries in eastern Canada culminated in
moratoria in all regions, beginning in the Maritime provinces (1984) and
following in Newfoundland (1992), Labrador (1998), and Québec (2000). Since
these closures, many populations have not increased as expected (REF): some 
have been assessed as considered threatened or endangered by the 
Committee on the Status of Endangered Wildlife in Canada \citep[COSEWIC, ][]{Cosewic2010}, 
Canada’s national science advisory body (to the national government) on species risk of extinction.
While it is not fully understood what is driving population declines, the potential
drivers of these are many \citep[see ][for a detailed discussion of possible
causes]{Cairns2001}: fishing mortality, damming of waterways
and changes in their freshwater habitats \citep{Dunfield1985}, acidification
of freshwater habitats \citep[particularly in the Southern Uplands region of
NS, see][]{Gibson2010}, predation by seals and birds \citep{Cairns2000}, negative
effects of interbreeding or interactions with escaped farmed salmon
\citep{Keyser2018}, and changes in their marine survival.

% 2.
Over the past three decades, a narrative has emerged that marine survival of Atlantic salmon has
declined throughout the North Atlantic \citep{Hansen1998,OMaoileidigh2003,Chaput2012a}. 
An implicit assumption, albeit infrequently acknowledged, is that any trend in
survival in ocean habitat that is shared by multiple populations during their
seaward migration period will be experienced similarly by multiple
populations. Put another way, given that salmon from different river systems
are hypothesized to share marine habitat during some of their time at sea, it
has been implicitly assumed that populations share similar temporal trends in
at-sea mortality \citep{Friedland1993, Friedland1998}.  
A recent study by \citet{Olmos2019} suggested that trends in post-smolt
survival, when estimated at the stock unit level, are synchronously declining
for all Atlantic salmon in eastern North America.

@@

In contrast to the narrative of widespread, demographically similar increases
in at-sea mortality, the conservation status of Canadian populations differs
considerably. Populations in the southern part of the Canadian range are more
likely to be assessed as being below conservation targets than those in more
northerly regions \citep{Cosewic2010}. This geographical disparity in status
suggests that if marine survival has been, or is, a key factor responsible for
most population declines, these changes are not uniformly distributed across
all populations.

@@

%4. 
In the present study, we compile smolt and adult returns times series data for seven wild
Canadian populations of Atlantic salmon to model trends in marine survival.
While some studies have used maturity-schedule models to estimate marine
survival for a limited number of salmon populations \citep{Chaput2003b}, none
have incorporated data extending over decates, nor have they examined
trends among more than two or three populations. 
Here, we develop a hierarchical Bayesian model that uses Murphy's maturity
schedule method, in conjunction with informative priors, to estimate yearly
marine survival in salmon. In addition to accounting for observation error in
smolt and return estimates, we estimate the proportion of salmon returning
after one winter hierarchically.

@@

We obtained time series data of outmigrating smolt and returning adult
abundances for seven Atlantic salmon populations in eastern Canada, encompassing a
wide range of the species' distribution (Fig.~\ref{fig:map}). These
populations included LaHave River in the Southern Uplands region of Nova
Scotia (NS), Nashwaak River in New Brunswick (NS) which connects to the
Bay of Fundy, Trinit\'{e} and Saint-Jean Rivers in Qu\'{e}bec (QC), which
flow to the St. Lawrence River and Bay of St. Lawrence, respectively, and
Western Arm Brook (also referred to as WAB), Campbellton, and Conne Rivers in Newfoundland, which flow to the
Strait of Belle Isle, Labrador Sea, and Gulf of St. Lawrence, respectively.
Data were collected in NS, NB, and Newfoundland
and Labrador (NL) by Fisheries and Oceans Canada (DFO) and in QC
by the Minist\`{e}re des For\^{e}ts, de la Faune et des Parcs, Qu\'{e}bec.

@@

\subsubsection*{Smolt data}

\comment{Co-authors: please check data sources}

Smolt estimates originate from a variety of sources.
Estimates from the XXXX populations were obtained using a capture-mark-recapture approach, while
Estimates from the XXXX populations were obtained by direct counts using a XXXX.
Refer to ... for more details on the data collection details.

\subsubsection*{Returns data}

@@

We developed a hierarchical Bayesian model that uses Murphy's maturity
schedule method, in conjunction with informative priors, to estimate yearly
marine survival in seven populations of Atlantic salmon. We account for
observation error in smolt and return estimates, as well as estimating the
proportion returning after one winter (i.e. \Pg) hierarchically.
There is an identifiability problem in the maturity schedule equations where
the parameter estimates cannot be optimally solved \citep{Chaput2003a}.
However, this issue can be mathematically overcome, at least partially, by
using informative priors for all three marine survival parameters in a
Bayesian framework.
This method requires abundance estimates of smolts as well as abundance estimates
of returning one-sea-winter (1SW) and two-sea-winter adults. With these data,
it estimates three parameters: survival in the first year at sea (\So), survival
in the second year at sea (\St), and the proportion of fish returning after one
year at sea (\Pg). 
Our model does not include repeat spawners and assumes that no fish spend
three or more winters at sea before returning to spawn for the first time.

The model also assumes that mortality in the second winter at sea 
is additional to mortality in the first winter at sea in the previous year, 
and therefore does not account for differences
the returning (Pr) occurs just before
actually being counted as returns and that S2 is any additional mortality in
the subsequent year. I agree this is far from ideal, given that we know there
is a period of a few months where 1SW returns are subject to a different
environment than those 2SW+ fishes.

@@

\begin{align}
\epsilon_{1,r} &\sim Normal(0, 0.01) \\
\epsilon_{2,r} &\sim Normal(0, 0.01)
\end{align}

@@

\begin{align}
 S_{1} &= e^{-Z_1}, \\
 S_{2} &= e^{-Z_2} 
\end{align}

@@

\begin{align}
    logit(P_{r,t}) &\sim Normal(\mu_r, \sigma_r) \\
    \mu_r &\sim 
    \begin{cases}
       Normal(2.3, 0.4),  &\text{for 1SW-dominated populations} \\
       Normal(0, 2.8), &\text{for non-1SW-dominated populations} \\
   \end{cases} \\
    \sigma_r &\sim halfNormal(0, 1)
\end{align}

@@

\begin{align}
Z_1 &\sim logNormal(1, 0.22),   \\ 
Z_2 &\sim logNormal(0.2, 0.3)
\end{align}

@@

\begin{figure}[htbp] \centering
    \includegraphics[width=0.95\linewidth]{figures/s1-trends-dual.png}
    \caption{Trends in survival in the first year at sea (\So). a) Posterior
        estimates of \So for the seven rivers examined, error bars indicate
        the 90\% credible intervals. b) Z-scores of median posterior
        estimates.} \label{fig:s1-dual} 
\end{figure}

@@

% 1. Main findings
Our findings challenge the narrative that marine survival is declining
uniformly throughout the range of Atlantic salmon in the northwest Atlantic.
Temporal trends are not consistent among populations 
. Over the time periods for which data were available, some rivers show positive trends in survival in the
first winter at sea (\So) while other show stable trends, and some even show
declines. We could not assess trends in the second winter at sea (\So) or
proportion returning as grilse (\Pg), as these parameter estimates were highly
uncertain and were strongly influenced by the priors used.
Perhaps there is a need to rethink our understanding of Atlantic salmon
population dynamics in light of the possibility that any real and persistent decline in marine survival was
experienced by some but not all populations, that reductions in survival might
have occurred over a relatively brief period of time and may not longer be ongoing, and that since that
reduction occurred (for a subset of populations) marine survival has
been stable, or even increasing.

@@

% 3. Potential reasons
There are multiple potential explanations for the lack of synchronous
trends in estimates of \So. 
Marine survival in the first winter at sea could be highly variable between
populations because of the predominance of spatially local environmental drivers of survival (e.g., temperature, predation) 
relative to broader-scale, even ocean-wide, drivers.
The synchrony reported for marine survival trends at broader spatial scales \citep{Olmos2019}
might be attributable to the use of stock-recruitment relationships to estimate survival,
relationships thay may have been confounded by changes in recruitment dynamics.

There is some evidence of a correlation between return rate and growth (as
indicated by intercirculi spacing on scales), where years of poor growth
tended to also be years of poor survival \citep{Friedland1993}, supporting the
idea that environmental variability can affect marine survival.
Furthermore, among European salmon, there is evidence of a positive correlation
between spring temperature in the Norwegian and North Seas and population abundance, suggesting warmer
conditions favour post-smolts \citep{Friedland1998} based on mapping the
extent of area of suitable temperature (7-13 \textdegree C).

Nonetheless, the causal mechanisms for why warming should affect post-smolt
survival almost certainly differs depending on the difference between
temperature experienced by the post-smolts and their respective
population-specific thermal optima, which could explain why populations in eastern North America are declining
in the southern part of their range but potentially increasing further north.

A potential association between temperature and survival warrants
further study.

@@

% 4. Caveats
As with all novel modelling approaches, there are caveats to acknowledge.
While there are analytical issues with estimation of \So, \St, and \Pg,
the assumption that \St is additive to \So could produce unrealistic results.
Overcoming this assumption would require an additional parameter to be
estimated, or an additional assumption as to what proportion of \So is not
additive to \St (as the returning 1SW adults do not experience the same
environment when they return to their natal streams as those fish who stayed
at sea for an additional winter before returning to spawn).
Secondly, the assumed hierarchical structure might not be the most appropriate
for modelling smolt abundances. This approach results in shrinkage to the
mean, which means that the variability of yearly smolt estimates is less than
it would be if not modelled hierarchically. 
That said, this assumption is likely to result in more conservative trends in marine survival, as the
variation is smolt estimates is reduced.
It is important to note that the hierarchical structure in the estimation of
\Pg seems like a reasonable assumption given that the probability of returning
as grilse has a genetic component associated to it \citep{Aykanat2019} and is
not expected to vary much, within a population, among years.

@@

% We would like to thank Geir Bolstad, G\'{e}rald Chaput, Brian Dempson, and Martha
% Robertson for their useful discussions on estimating marine survival in
% Atlantic salmon, and Amanda Kissel for her helpful comments on the manuscript.
% Brian Dempson and Geoff Venoit for providing the data Conne River data.
We would like to thank G\'{e}rald Chaput for his useful discussions on
estimating marine survival in Atlantic salmon, and Sean Anderson for his help
with implementing the non-centered parameterization of the model.
This research was supported by the Atlantic Salmon Conservation Foundation and
the Atlantic Salmon Research Joint Venture.

>>>>

\usepackage[round,sectionbib]{natbib}
\setcitestyle{authoryear}
\bibpunct{(}{)}{;}{a}{}{;}
\bibliographystyle{fishfishnourl}

@@

\noindent\small{\textsuperscript{1} Department of Biology, Dalhousie University, Halifax, NS, B3H 4R2, Canada\\}
\small{\textsuperscript{2} Norwegian Institute for Nature Research (NINA), Trondheim, Norway\\}
\small{\textsuperscript{3} Fisheries and Oceans Canada, St. John's, NL, Canada\\}
\small{\textsuperscript{4} Minist\`{e}re des For\^{e}ts, de la Faune et des Parcs, Qu\'{e}bec, QC, Canada\\}
\small{\textsuperscript{5} Fisheries and Oceans Canada, Moncton, NB, Canada\\}
\small{\textsuperscript{6} Fisheries and Oceans Canada, Dartmouth, NS, Canada\\}
\small{\textsuperscript{*} Corresponding author: spardo@dal.ca}

@@

The marine phase of anadromous Atlantic salmon (\emph{Salmo salar}) is the
least known yet one of the most crucial with regards to population
persistence. Declines in many Atlantic salmon populations in eastern
Canada have often been attributed to changes in conditions at sea, negatively
affecting their survival. However, marine survival estimates are difficult to
obtain given that many individuals spend multiple winters in the ocean before
returning to freshwater to spawn, necessitating the estimation of multiple
parameters. To do so, we fit a hierarchical Bayesian maturity schedule model
to smolt and adult abundance time series for seven populations located in
Newfoundland and Labrador (NL), Qu\'{e}bec (QC), New Brunswick (NB), and Nova
Scotia (NS). The datasets ranged between 19 and 42 years in length. We
estimated three components of marine survival: survival in the first year at
sea (\So), survival in the second year at sea (\St), and proportion returning
as one sea-winter adults (\Pg). Controlling for time frame, trends in
estimates of \So were not consistent among rivers.
In the four populations for which data extended to 1990, marine survival
during the first year at sea predominantly increased over time in Western Arm
Brook (NL) and Rivi\`{e}re Saint-Jean (QC) but largely declined in Conne River
(NL) and Rivi\`{e}re Trinit\'{e} (QC). In the three other populations, \So
exhibited an increase in Campbellton River (NL), a decline in LaHave River
(NS), and fluctuating stability in Nashwaak River (NB) since approximately the
mid-1990s. 
Estimates of \St were highly uncertain, particularly for 1SW-dominated rivers
(Conne, Campbellton and WAB, where the abundance of 2SW returns is very low)
and thus the posterior distributions matched our choice of prior and trends
could not be assessed. 
\Pg was temporally stable within rivers, except for LaHave and Nashwaak Rivers. 
Our findings challenge the assumption that marine survival has changed in a consistent
manner among populations across a broad geographic scale, and suggest that
trends can be river-specific. 
The well established correlations between climate variables and abundance are
not being mediated solely by marine survival, and there can potentially be
some important indirect effects on fecundity.
Further work exploring potential correlates of marine survival and the
potential non-lethal effects of climate effects is warranted.

@@

Reductions in fishing mortality, albeit necessary, are not always sufficient
to facilitate population recovery. Experience with numerous commercially
exploited marine fisheries since the early 1990s has shown that not all
populations respond as favourably as anticipated to considerable reductions in
exploitation \citep{Hutchings2017}. Efforts to close commercial
Atlantic salmon (\emph{Salmo salar}) fisheries in eastern Canada culminated in
moratoria in all regions, beginning in the Maritime provinces (1984) and
following in Newfoundland (1992), Labrador (1998), and Qu\'{e}bec (2000). Since
these closures, many populations have not increased as 
expected \citep{Dempson2004, ICES2019}; some 
have been assessed as considered threatened or endangered by the 
Committee on the Status of Endangered Wildlife in Canada \citep[COSEWIC, ][]{Cosewic2010}, 
Canada's national science advisory body (to the national government) on
species risk of extinction.
While it is not fully understood what is driving population declines, the potential
drivers of these are many \citep[see ][for a detailed discussion of possible
causes]{Cairns2001}, including but not limited to: fishing mortality \citep{Dempson2004}, 
damming of waterways and changes in their freshwater habitats \citep{Dunfield1985}, acidification
of freshwater habitats \citep[particularly in the Southern Uplands region of
NS, see][]{Gibson2010}, predation by seals and birds \citep{Cairns2000}, negative
effects of interbreeding or interactions with escaped farmed salmon
\citep{Keyser2018}, and climate-driven changes in survival and productivity \citep{Mills2013}.

% 2.
Over the past three decades, a narrative has emerged that marine survival of
Atlantic salmon has declined throughout the North Atlantic \citep{ICES2019}.
%\citep{Hansen1998,OMaoileidigh2003,Chaput2012a}.
As there are multiple lines of evidence that climate conditions directly and
indirectly influence the abundance and productivity of Atlantic salmon
populations \citep{Mills2013,Almodovar2019}, it is expected that these oceanic climate effects are mostly
driving population dynamics through changes in marine survival.
An implicit assumption, albeit infrequently acknowledged, is that any trend in
survival in ocean habitat that is shared by multiple populations during their
seaward migration period will be experienced similarly by multiple
populations. Put another way, given that salmon from different river systems
are hypothesized to share marine habitat during some of their time at sea, it
has been implicitly assumed that populations share similar temporal trends in
at-sea mortality \citep{Friedland1993, Friedland1998, Russell2012}.  
A recent study by \citet{Olmos2019} suggested that trends in post-smolt
survival, when estimated at the stock unit level, are synchronously declining
for all Atlantic salmon in eastern North America.

@@

In contrast to the narrative of widespread, demographically similar increases
in at-sea mortality, the conservation status of Canadian populations differs
considerably. Populations in the southern part of the Canadian range are more
likely to be assessed as being of conservation concern than those in more
northerly regions \citep{Cosewic2010}. This geographical disparity in status
suggests that if marine survival has been, or is, a key factor responsible for
most population declines, these changes are not uniformly distributed across
all populations.

@@

%4. 
In the present study, we compile smolt and adult returns times series data for seven wild
Canadian populations of Atlantic salmon to model trends in marine survival.
While some studies have previously used maturity-schedule models to estimate marine
survival for a limited number of salmon populations \citep{Chaput2003b}, none
have incorporated data extending over decades, nor have they examined
trends among more than two or three populations. 
Here, we develop a hierarchical Bayesian model that uses Murphy's maturity
schedule method, in conjunction with informative priors, to estimate yearly
marine survival in salmon. In addition to accounting for observation error in
smolt and return estimates, we estimate the proportion of salmon returning
after one winter hierarchically.

@@

We obtained time series data of outmigrating smolt and returning adult
abundances for seven Atlantic salmon populations in eastern Canada, encompassing a
wide range of the species' distribution (Fig.~\ref{fig:map}). 
Populations included the LaHave River in the Southern Uplands region of Nova
Scotia (NS), Nashwaak River, New Brunswick (NB), Rivi\`{e}re de la Trinit\'{e} and
Saint-Jean rivers in Qu\'{e}bec (QC), and  Western Arm Brook (also referred to asalso referred to as  WAB), Campbellton, and
Conne River, Newfoundland (NL). 
Data were collected in NS, NB, and Newfoundland
and Labrador (NL) by Fisheries and Oceans Canada (DFO) and in QC
by the Minist\`{e}re des For\^{e}ts, de la Faune et des Parcs, Qu\'{e}bec.

@@

\subsubsection*{Smolt and adult return abundance data}

Smolt and adult return abundance estimates originate from a variety of sources.
Estimates from the XXXX populations were obtained using a capture-mark-recapture approach, while
Estimates from the XXXX populations were obtained by direct counts using a XXXX.
... for more details on the . 
For data collection details refer to 
\citet{Venoitt2018} for NL populations, 
\citet{April2018}  for QC populations,
\citet{Jones2014} for NB populations,
and \textbf{REF} for NS populations.

@@

We developed a hierarchical Bayesian model that uses Murphy's maturity
schedule method, in conjunction with informative priors, to estimate yearly
marine survival in seven populations of Atlantic salmon. We account for
observation error in smolt and return estimates, as well as estimating the
proportion returning after one winter (i.e. \Pg) hierarchically.
There is an identifiability problem in the maturity schedule equations where
the parameter estimates cannot be optimally solved \citep{Chaput2003a}.
However, this issue can be mathematically overcome, at least partially, by
using informative priors for all three marine survival parameters in a
Bayesian framework.
This method requires abundance estimates of smolts as well as abundance estimates
of returning one-sea-winter (1SW) and two-sea-winter adults. With these data,
it estimates three parameters: survival in the first year at sea (\So), survival
in the second year at sea (\St), and the proportion of fish returning after one
year at sea (\Pg).

Our model does not include repeat spawners and assumes that no fish spend
three or more winters at sea before returning to spawn for the first time.
The model also assumes that mortality in the second winter at sea (\St)
is additional to mortality in the first winter at sea in the previous year, 
and therefore does not account for differences in environmental conditions experienced
between 1SW and 2SW fish of the same smolt cohort during their overlapping first year at sea.
In other words, our model assumes that the decision of returning occurs just before
actually being counted as returns and that \St is any additional mortality in
the subsequent year.

@@

\begin{align}
\epsilon_{1,r} &\sim Normal(0, 0.01) \\
\epsilon_{2,r} &\sim Normal(0, 0.01).
\end{align}

@@

\begin{align}
 S_{1} &= e^{-Z_1}, \\
 S_{2} &= e^{-Z_2}. 
\end{align}

@@

\begin{align}
    logit(P_{r,t}) &\sim Normal(\mu_r, \sigma_r) \\
    \mu_r &\sim 
    \begin{cases}
       Normal(2.3, 0.4),  &\text{for 1SW-dominated populations} \\
       Normal(0, 2.8), &\text{for non-1SW-dominated populations} \\
   \end{cases} \\
    \sigma_r &\sim halfNormal(0, 1).
\end{align}

@@

\begin{align}
Z_1 &\sim logNormal(1, 0.22),   \\ 
Z_2 &\sim logNormal(0.2, 0.3).
\end{align}

@@

\begin{figure}[htbp] \centering
    \includegraphics[width=0.95\linewidth]{figures/s1-trends-dual.png}
    \caption{Trends in survival in the first year at sea (\So). a) Posterior
        estimates of \So for the seven rivers examined, error bars indicate
        the 90\% credible intervals, and dashed lines denote years of commercial
        fishing moratoria for each province. b) Z-scores of median posterior
        estimates.} \label{fig:s1-dual} \end{figure}

@@

% 1. Main findings
Our findings challenge the narrative that marine survival, specifically survival in the first year at sea, is declining
uniformly throughout the range of Atlantic salmon in the northwest Atlantic.
Temporal trends are not consistent among populations 
. Over the time periods for which data were available, some rivers show positive trends in survival in the
first winter at sea (\So) while other show highly variable yet stable trends, and some even show
declines. We could not assess trends in the second winter at sea (\So) or
proportion returning as grilse (\Pg), as these parameter estimates were highly
uncertain and were strongly influenced by the priors used.
Perhaps there is a need to rethink our understanding of Atlantic salmon
population dynamics in light of the possibility that any real and persistent decline in marine survival was
experienced by some but not all populations, that reductions in survival might
have occurred over a relatively brief period of time and may not longer be ongoing, and that since that
reduction occurred (for a subset of populations) marine survival has
been stable, or even increasing.

@@

% 3. Potential reasons
There are multiple potential explanations for the lack of synchronous
trends in estimates of \So. 
Marine survival in the first winter at sea could be highly variable between
populations because of the predominance of spatially local environmental drivers of survival (e.g., temperature, predation) 
relative to broader-scale, even ocean-wide, drivers.
The synchrony reported for marine survival trends at broader spatial scales \citep{Olmos2019}
might be attributable to the use of stock-recruitment relationships to estimate survival,
relationships that may have been confounded by changes in recruitment dynamics.
There is some evidence of a correlation between return rate and growth (as
indicated by intercirculi spacing on scales), where years of poor growth
tended to also be years of poor survival \citep{Friedland1993}, supporting the
idea that environmental variability can affect marine survival.
Furthermore, among European salmon, there is evidence of a positive correlation
between spring temperature in the Norwegian and North Seas and population abundance, suggesting warmer
conditions favour post-smolts \citep{Friedland1998} based on mapping the
extent of area of suitable temperature (7-13 \textdegree C).

Nonetheless, the causal mechanisms for why warming should affect post-smolt
survival almost certainly differs depending on the difference between
temperature experienced by the post-smolts and their respective
population-specific thermal optima. 
This difference could explain why populations in eastern North America are
declining in the southern part of their range but potentially increasing
further north, and also why some studies find possitive correlations between
temprature and abundance \citep{Friedland1998, Friedland1998b, Jonsson2004}
while other find negative ones \citep{Friedland1993, Todd2008}.
A potential association between temperature and direct estimates of marine
survival warrants further study.

@@

Oceanic conditions have been correlated with abundance trends and growth
\citep{Todd2008}, however, the mechanism by which bottom-um effects, mediated by changes in food availability,
population dynamics beyond marine survival needs to
be thoroughly reassessed. If marine survival on its own cannot fully explain
trends in abundance, then there are potential carry-on effects of oceanic
conditions that manifest with regards to fresh production. 
For example, adults
that return to spawn after spending suboptimal conditions at sea might be less
likely to make it to their spawning grounds, successfully secure a mate,
produce fewer eggs, or produce eggs with lower per capita fitness than those
produced by adults which reared in optimal oceanic conditions.
As larger females tend to be more productive, in terms of fecundity and total
reproductive energy, than the same weight's worth of smaller females
\citep{Barneche2018}, a small decrease in body condition resulting from bottom-up
impacts on food availability, could potentially have disproportionate effects on fecundity
and fitness of the offspring.
Egg-to-smolt survival in Atlantic salmon is highly variable \citep{Chaput2015}
and changes in the oceanic conditions that spawners experiences could be
contributing to this variability.
Obviously, there would be a delay in these effects to manifest as adult
trends, as it would take a whole life cycle for fecundity effects to manifest
as abundance of returning adults.
However, given that most correlations are between relatively monotonic declines
in abundance coupled also monotonic increases in climatic indices
over decadal time scales \citep[e.g.,][]{Friedland1998, Todd2008,
    Beaugrand2012}, it would be expected that this correlations would be
maintained even if salmon abundances were lagged by a generation length.

% 4. Caveats
As with all novel modelling approaches, there are caveats to acknowledge.
The seven populations explored in this study might not be representative
of the regional trends in marine survival. However, there are no other
long-term time series of smolts and adult returns to draw inferences from.
While there are analytical issues with estimation of \So, \St, and \Pg,
the assumption that \St is additive to \So could produce unrealistic results.
We know there is a period of a few months where 1SW
returns are subject to a different environment than those salmon that will
return as 2SW the next year. 
While this is not ideal,
overcoming this assumption would require an additional parameter to be
estimated, or an additional assumption as to what proportion of \So is not
additive to \St (as the returning 1SW adults do not experience the same
environment when they return to their natal streams as those fish who stayed
at sea for an additional winter before returning to spawn).
Secondly, the assumed hierarchical structure might not be the most appropriate
for modelling smolt abundances. This approach results in shrinkage to the
mean, which means that the variability of yearly smolt estimates is less than
it would be if not modelled hierarchically. 
That said, this assumption is likely to result in more conservative trends in marine survival, as the
variation is smolt estimates is reduced.
It is important to note that the hierarchical structure in the estimation of
\Pg seems like a reasonable assumption given that the probability of returning
as grilse has a genetic component associated to it \citep{Aykanat2019} and is
not expected to vary much, within a population, among years.

@@

% We would like to thank Geir Bolstad, G\'{e}rald Chaput, Brian Dempson, and
% Martha Robertson for their useful discussions on estimating marine survival
% in Atlantic salmon, and Amanda Kissel for her helpful comments on the
% manuscript. Brian Dempson and Geoff Venoit for providing the data Conne
% River data.
We would like to thank G\'{e}rald Chaput for his useful discussions on
estimating marine survival in Atlantic salmon, Carmen David for her comments
on the manuscript, and Sean Anderson for his help with implementing the
non-centered parameterization of the model. This research was supported by the
Atlantic Salmon Conservation Foundation and the Atlantic Salmon Research Joint
Venture.
